---
title: "STA 308 -- Final Exam"
author: "Matt Fry"
date: "`r Sys.Date()`"
output: html_document
---

```{r}
library(tidyverse)
```


# Data Task
## The file gasoline_prices.csv Download gasoline_prices.csvcontains the state 
## average gasoline prices for each state (and Washington, DC) in the United 
## States on November 29, 2022 according to AAA. The file state_abb_codes.csv 
## state_abb_codes.csvprovides a mapping of states & districts in the United 
## States with their abbreviations, and the file censusRegions.csv 
## censusRegions.csvprovides a mapping of states and districts based on 
## census regions. 

## For each of the census regions, compute the mean state (that is, remove 
## Washington, D.C.) percentage increase (difference) in the price of Diesel 
## vs Regular gasoline ( (Diesel - Regular)/Regular*100), along with the 
## corresponding standard deviation and coefficient of variation . The result 
## of your program should be a table of numbers with 4 rows (corresponding 
## to the four census regions) and 3 columns of calculated values (the mean, 
## standard deviation and coefficient of variation) [note: your output may 
## be 4 columns if you include/consider census regions).

# Takeaways from instructions:
## Compute the mean percentage increase in the price of Regular vs Diesel by state
## Remove Washington DC
## Group states by census region
## Also compute standard deviation and coefficient of variation

## Website for fuel prices by state: "https://tjfisher19.github.io/data/gasoline_prices.csv"
## Website for state abbreviations: "https://tjfisher19.github.io/data/state_abb_codes.csv"
## Website for census regions: "https://tjfisher19.github.io/data/censusRegions.csv"

```{r}
## Reading in all three data sets
FuelPriceStateDF <- read_csv("https://tjfisher19.github.io/data/gasoline_prices.csv")
StateAbbrevsDF <- read_csv("https://tjfisher19.github.io/data/state_abb_codes.csv")
CensusRegionsDF <- read_csv("https://tjfisher19.github.io/data/censusRegions.csv")
```


```{r}
## Now that I have read in the data, I am going to select the data I need to 
##  reduce the unneccessary information in my final data frame and find the 
##  price difference between Diesel and regular
FuelPriceStateDF <- FuelPriceStateDF %>%
  mutate(Pct_PriceDiff = (Diesel - Regular)/Regular*100) %>%
  select(State, Pct_PriceDiff) %>%
  filter(State != "District of Columbia") ## removing Washington DC
StateAbbrevsDF <- StateAbbrevsDF %>%
  select(State, Code)
```

```{r}
## Now that I have the data I NEED, I am going to merge datasets to have the 
##  census regions and provide summary statistics by region

StatesWithCodes <- merge(FuelPriceStateDF, StateAbbrevsDF) %>%
  select(Code, Pct_PriceDiff)
PriceDiffWithRegion <- merge(StatesWithCodes, CensusRegionsDF, 
                           by.x = "Code", by.y = "State")
## Now that the datasets are merged, I need to group the states by region
##  in order to find the average percent difference in the cost of 
##  Diesel vs Regular by region, as well as the standard deviation and 
##  coefficient of variation

StatsByRegion <- PriceDiffWithRegion %>%
  group_by(Region) %>%
  summarize(RegionalAvg = mean(Pct_PriceDiff), 
            RegionalStdDev = sd(Pct_PriceDiff), 
            RegionalCoeffOfVar = RegionalStdDev/RegionalAvg)
print(StatsByRegion)
## "RegionalAvg" is the average of the percent difference between Diesel and 
## Regular by region
## "RegionalStdDev" is the standard deviation of the percent difference 
## by region
## "RegionalCoeffOfVar" is the coefficient of variation by region --
## coefficient of variation is the standard deviation/mean

```








